Precision calculations of the cosmic shear power spectrum projection

### 2017-02-16 limber2.tex
Added refs for DLS and DES-SV to add to Table 1 Some minor edits Anything more significant is in purple One TODO martin - add a ref for the Legendre equation

We further show that the second-order Limber approximation is an
accurate representation of the full projection, with better than percent level
precision for scales $\ell > 3$. Since this approximation involves only 1D
integrals over the matter power spectrum, it is very fast to calculate
numerically and can readily be employed in Monte-Carlo sampling methods to
obtain precision constraints on cosmological parameters.

\mk{We also compute the shear correlation function using a spherical Legendre
transformation, and compare this to the flat-sky approximated, commonly used
fast Hankel transformation.}

This paper is organised as follows.  In section~\ref{sec:wl} we provide a
pedagogical introduction to weak gravitational lensing theory, projections and
power spectra for the flat-sky and spherical cases, following the seminal work
from \citet{2000PhRvD..62d3007H}, see also \citet{2005PhRvD..72b3516C}. In
section~\ref{sec:L2} we derive weak-lensing observables using a second-order
Limber approximation first introduced by \citet{2008PhRvD..78l3506L}. We
compare the shear power spectrum and the commonly-used two-point shear
correlation function for the full solution to a range of different
approximations in section~\ref{sec:results}, providing cosmological constraints
for each case using CFHTLenS data from \citet{CFHTLenS-2pt-notomo}. This paper
draws from several sources of literature which have previously discussed the
full projection, or reviewed the accuracy of the Limber approximation for weak
lensing studies, namely \citet{2008PhRvD..78d3002S,2012PhRvD..86b3001B,
2012MNRAS.422.2854G, 2016arXiv161104954K}. We present a discussion and
comparison of our results to these papers in Appendix~\ref{app:B}.

@@

\mk{This type of cross-power spectrum between different cosmological epochs
$\chi$ and $\chi^\prime$ was introduced in \cite{2005PhRvD..72b3516C}. In
Sects.~\ref{sec:geom_mean} and \ref{sec:factor_D} this unequal-time
cross-spectrum \citep{2016arXiv161200770K} will be further evaluated and
simplified in the context of the Limber approximation. The oscillating Bessel
functions  in (\ref{eq:C_ell_phi_Pm}) ensure that only relatively close epochs
contribute to the lensing potential correlation. This make sense since observed
light rays from two galaxies at different positions on the sky that necessarily
converge at the observer today, pick up the density fluctuations at similar
times while propagating through the large-scale structure. A similar argument
has been made in \cite{BS01}: since the matter power spectrum scales with $k$
for $k \rightarrow 0$, there is decreasing power towards larger and larger
scales. In particular, the correlation of cosmic fields decreases strongly
above a coherence scale $|\chi - \chi^\prime| \gsim L_{\rm coh}$ which is much
smaller than the horizon scale $\chi_{\rm h}$. }

@@

The majority of cosmic shear analyses have used the predicted lensing power
spectrum approximated on a flat sky, neglecting the sky curvature. This is a
valid approach for past and current survey areas with an extent less than 10
degrees. To account for the sky curvature of the observed data, the shear
correlation functions from observed galaxy ellipticities is now routinely
computed using spherical coordinates, since projecting to a Cartesian plane has
been shown to cause significant biases of the two-point funtion on large scales
\citep{FSHK08}, \mk{and lead to spurious B-modes \cite{asgari/etal:2017}}. Here
we examine the effect of sky curvature on the theoretical models of the power
spectrum, and the effect on cosmological parameter inference (see
Sect.~\ref{sec:cfhtlens}).

@@

To separate the $k$- and $\chi, \chi^\prime$-terms in
equation~(\ref{eq:C_ell_full}), we first approximate the matter power
cross-spectrum between two distances by the geometric mean of the two involved
distances \citep{2005PhRvD..72b3516C,2016arXiv161200770K},
%
\begin{equation}
 P_{\rm m}(k; \chi, \chi^\prime) = \sqrt{ P_{\rm m}(k; \chi) P_{\rm m}(k; \chi^\prime) } \, .
  \label{eq:P_geom_mean}
\end{equation}
%
\mk{This form is justified when considering the linear power spectrum, and
follows when inserting the linearly evolving density contrast $\hat \delta(\vec
k; \chi) = D_+(\chi) \hat \delta_0(\vec k)$ into (\ref{eq:p_m}), whhere
$\delta_0$ is the present-day linearly extrapolated density contrast, and $D_+$
the linear growth factor with $D_+(0) = 1$. This is a very good approximation
also in the non-linear case as shown in \cite{2016arXiv161200770K}.}

@@

Following \cite{2008PhRvD..78l3506L} we expand to third order
%
\begin{equation}
  \lim_{\varepsilon \rightarrow 0} \int_0^\infty {\rm d} x \, {\rm e}^{-\epsilon (x - \nu)} g(x) {\rm J}_\nu(x)
  %\approx g(\nu) - \frac 1 2 g^{\prime\prime}(\nu) - \frac \nu 6 g^{\prime\prime\prime}(\nu) 
  \approx g(\nu) - \frac 1 2 \left.\frac{{\rm d}^2 g}{{\rm d} x^2}\right|_{x=\nu}
                 - \frac \nu 6 \left.\frac{{\rm d}^3 g}{{\rm d} x^3}\right|_{x=\nu},
\end{equation}
%
with $g(x) = k^{-1} f(k, \chi)$, $\chi=x/k$, and its derivatives $g^{(n)}(x) =
k^{-1-n} f^{(n)}(k, \chi)$ for a given $k$, where the derivatives of $f$
are with respect to the second argument $\chi$. \mk{This series expansion, the
replacement of the integral with the evaluation of $g$ and its derivatives at
the maximum of the Bessel function is a good approximation of the integral if
$g$ is varying more slowly than the oscillating Bessel function.
% [Check this].
As we will see below, $f$ is a slowly varying function of the comoving
distance.} In our case the projection kernel is
%
\begin{equation}
  f(k, \chi) = \sqrt{P_{\rm m}(k; \chi)} \, a^{-1}(\chi) \chi^{-3/2} q(\chi).
  \label{eq:f_LA08}
\end{equation}
%
In the tomographic case, the index $i$ is added to $q$ and $f$.
%
Replacing both distance integrals in equation~(\ref{eq:C_ell_full_Pk}) by their
Taylor-expansions around the maxima $\nu(\ell) = \ell + 1/2$ of the two Bessel
functions, which are $k \chi$ and $k \chi^\prime$, respectively, yields
%
\begin{align}
  C^\gamma_{ij}(\ell) \approx & \, \ellbar^2(\ell, 2) \, \pref^2
    \int_0^\infty \frac{{\rm d} k}{k^3} \, k^{-2}
    \left[ f_i(k, \chi) - \frac{1}{2 k^2} f_i^{\prime\prime}(k, \chi)
      - \frac{\nu(\ell)}{6 k^3} f_i^{\prime\prime\prime}(k, \chi) + \ldots \right]
    %\nonumber \\
    %& \times
    \left[ f_j(k, \chi) - \frac{1}{2 k^2} f_j^{\prime\prime}(k, \chi)
    - \frac{\nu(\ell)}{6 k^3} f_j^{\prime\prime\prime}(k, \chi) + \ldots \right] \, .
  \label{eq:C_ell_limber2_dk}
\end{align}
%
Changing the integration to $\chi = \nu/k$ and collecting terms according to their $\nu$-dependence:
%
\begin{align}
  C^\gamma_{ij}(\ell) \approx & \, C^\gamma_{{\rm L1}, ij}(\ell) + C^\gamma_{{\rm L2}, ij}(\ell)
      %\nonumber \\
    = %&
    \frac{\ellbar^2(\ell, 2)}{\nu^4(\ell)} \, \pref^2
    \int_0^\infty {\rm d} \chi \, \chi^3 \, % P_{\rm m}\left( k \right)
    \Bigg\{
    %\left\{
    (f_i f_j)(\nu(\ell)/\chi, \chi)
    %\right.
      \nonumber \\
    &
    %\left.
     - \frac 1 {\nu^2(\ell)}
    \left[ \frac{\chi^2}{2} \left( f^{}_i f^{\prime\prime}_j + f_i^{\prime\prime} f^{}_j \right)(\nu(\ell)/\chi, \chi)
     + \frac{\chi^3}{6} \left( f^{}_i f^{\prime\prime\prime}_j + f^{\prime\prime\prime}_i f^{}_j \right)(\nu(\ell)/\chi, \chi)
    \right]
    + {\cal O}(\nu^{-4})
    %\right\}
    \Bigg\}
    \, .
  \label{eq:C_ell_limber12_dr}
\end{align}
%
The first term corresponds to the well-known first-order Limber approximation
\citep{1953ApJ...117..134L,1992ApJ...388..272K}, which is widely used in
weak gravitational lensing. We retrieve the (spherical) standard expression by
inserting back the projection kernel (equation~\ref{eq:f_LA08}),
%
\begin{align}
  C^\gamma_{{\rm L1}, ij}(\ell) = & \, \frac{\ellbar^2(\ell, 2)}{\nu^4(\ell)} \, \pref^2 \int {\rm d} \chi \frac{ q_i(\chi) q_j(\chi) }{a^2(\chi)}
  P_{\rm m}\left(\frac{\nu(\ell)}{\chi}; \chi\right) . \Label{ExtL1Sph}
  \label{eq:C_ell_limber1}
\end{align}
%
In the Limber approximation, modes between structures at different epochs do
not contribute to the single line-of-sight integration.

@@

With the definition (\ref{eq:P_k_chi_sep_s}), we factor out the function
$\tilde P_{\rm m}(k, \chi)$ from equation~(\ref{eq:f_LA08}), and we define the
separated kernel function $f_{\rm s}$ as
%
\begin{equation}
  f_{\rm s}(\chi) = D_+(\chi) a^{-1}(\chi) \chi^{-3/2} q(\chi) \,
  \label{eq:f_LA08_s}
\end{equation}
%
The \mk{Limber equation up to second order} of the shear power spectrum can then be approximated as
%
\begin{align}
  C^\gamma_{ij}(\ell) = & \, C^\gamma_{{\rm L1}, ij}(\ell) 
    \nonumber \\
    & - \frac 1 {\nu^2(\ell)} \, \frac{\ellbar^2(\ell, 2)}{\nu^4(\ell)} \,
    \frac{\pref^2}{2}
    \int {\rm d} \chi \, \chi^{7/2} a^{-1}(\chi) D_+^{-1}(\chi) P_{\rm m}\left(\frac{\nu(\ell)}\chi; \chi\right)
    %\nonumber \\
    %& \times
    \left[ q^{}_i f^{\prime\prime}_{{\rm s} j} + f^{\prime\prime}_{{\rm s}i} q^{}_j
      + \frac{\chi}{3} \left( q^{}_i f^{\prime\prime\prime}_{{\rm s}j} + f^{\prime\prime\prime}_{{\rm s}i} q^{}_j
      \right)
    \right](\chi) \, . \Label{ExtL2Sph}
  \label{eq:C_ell_limber2_dr_s}
\end{align}
%
We compute the numerical higher derivatives as follows. The functions $f_{\rm
s}(\chi)$ are fitted as power laws with index $\approx \mk{-1.5}$, \mk{which is
expected if the growth suppression factor $D_+(a)/a$ varies slowly with $\chi$,
and the lensing efficiency $q \approx 1$ for small and medium $\chi$ given the
CFHTLenS redshift range.} The fit is performed between $\chi_{\rm min} = 0.001
\, {\rm Mpc}/h$ and $\chi_{\rm max} = 500 \, \mbox{Mpc}/h$. At larger comoving
distances the kernel decreases faster than a power law, so we exclude this
range from the fit. Even though on those scales the derivatives are larger due
to the steeper decline, the associated errors are very small as these scales
are down-weighed by the kernel function $f_{\rm s}$ itself. At $\chi = 500 \,
\mbox{Mpc}/h$ the filter function is less than $10^{-4}$ of its value 
at $1$\,Mpc$/h$.

@@

The extended flat-sky Limber approximation is readily derived from the
spherical case (\ref{eq:C_ell_limber1}, \ref{eq:C_ell_limber2_dr_s}), by
replacing the pre-factor $\ellbar^2(\ell, 2)$ with $\ell^4$,
%
\begin{align}
  P^\gamma_{{\rm L1}, ij}(\ell) = & \, p(\ell) \, \pref^2 \int {\rm d} \chi \frac{ q_i(\chi) q_j(\chi) }{a^2(\chi)}
  P_{\rm m}\left(\frac{\nu(\ell)}{\chi}; \chi\right);
  \label{eq:P_ell_limber1}
  \\
    P^\gamma_{{\rm L2}, ij}(\ell) = & - \frac 1 {\nu^2(\ell)} \, p(\ell) \, \frac{\pref^2}{2}
    \int {\rm d} \chi \, \chi^{7/2} a^{-1}(\chi) D_+^{-1}(\chi) P_{\rm m}\left(\frac{\nu(\ell)}\chi; \chi\right)
    %\nonumber \\
    %& \times
    \left[ q^{}_i f^{\prime\prime}_{{\rm s}j} + f^{\prime\prime}_{{\rm s}i} q^{}_j  
      + \frac{\chi}{3} \left( q^{}_i f^{\prime\prime\prime}_{{\rm s}j} + f^{\prime\prime\prime}_{{\rm s}i} q^{}_j
      \right)
    \right](\chi)
  \label{eq:P_ell_limber2}
\end{align}
%
Further approximations can be made for the pre-factor $p(\ell) = \ell^4/\nu^4(\ell)$ and $\nu(\ell)$:
%
\begin{enumerate}
  \item $p(\ell) = 1$, this corresponds to $\nu(\ell) = \ell$, which is the
    \emph{standard} Limber approximation \LabelTxt{L1Fl}
    Until recently, i.e.~for all pre-2014 CFHTLenS results, this was the approximation of choice.
    Note that we do not discuss the second-order Limber approximation \mk{with this $p$} here.
  %
  \item $p(\ell) = \ell^4/(\ell + 1/2)^4$. This corresponds to the \emph{extended} Limber
    approximation \LabelTxt{ExtL1Fl,ExtL2Fl} with $\nu(\ell) = \ell + 1/2$; however the following case is typically employed:
  %
  \item $p(\ell) = 1$, but keeping as argument of the power spectrum $\nu(\ell) = \ell + 1/2$. This is
    a \emph{hybrid} between
    standard and extended Limber approximation \Label{ExtL1FlHyb,ExtL2FlHyb}, and the first-order case 
    was used in \cite{KiDS-450,joudaki/etal:2016,joudaki/etal:2017}. As is shown
    below, this is a better approximation to the full projection than case (ii). In (\ref{eq:P_ell_limber2})
    the second-order suppression factor is also left to be $\nu^{-2}(\ell) = (\ell + 1/2)^{-2}$, providing
    a slightly more accurate approximation compared to $\nu^{-2}(\ell) = \ell^{-2}$.
\end{enumerate}
%

@@

Considering first the flat-sky cases, the standard first-order Limber
approximation, (L1Fl), that was adopted for all pre-2014 CFHTLenS analyses, we
find \mk{it} to be accurate to better than 10\% for $\ell>3$, converging slowly to the
true projection with percent level precison at $\ell>100$. For the extended
Limber approximations `hybrid' cases (ExtL1FlHyb and ExtL2FlHyb), despite
decreased accuracy for $\ell < 8$ in comparison to the standard first-order
Limber case, the errors with respect to the true power spectrum decrease much
faster, as $\ell^{-2}$, such that percent-level precision is reached at
$\ell>15$.  The first-order extended Limber approximation `hybrid' cases
(ExtL1FlHyb) was adopted by \citet{joudaki/etal:2016}, \citet{joudaki/etal:2017}, and
\cite{KiDS-450}\footnote{We confirm that there is a typographical error in
equation 4 of \cite{KiDS-450} which should include the extra term of `+0.5' in
$\nu(\ell)$ that was incorporated in the cosmological analysis.}.

@@

On a sphere, correlation functions formally cannot be related to the power
spectrum by the Hankel transform in equation~(\ref{eqn:xiGG}), and should
be replaced by a Legendre transform,
%
\mk{
\begin{equation}
 \xi_+(\theta) = \frac 1{4\pi} \sum\limits_{\ell=2}^\infty (2 \ell + 1) C^\gamma(\ell) {\rm P}_\ell[\cos(\theta)],
  \label{eq:xi_Legendre}
\end{equation}
%
where ${\rm P}_\ell(x)$ are the Legendre polynomials in $x$ of order $\ell$.}

The spherical power spectrum is formally not defined for non-integer $\ell$
\citep[see][for alternative spherical-sky formulae for the two-point
correlation function]{2005PhRvD..72b3516C}, as functions defined on the
sphere are necessarily periodic.
As we have shown in section~\ref{sec:comp}, however, the spherical second-order
extended Limber approximation provides a percent-level precision representation
of the full spherical projection for $\ell > 3$.
The spherical pre-factor $\ellbar(\ell, 2)$ (equation~\ref{eq:ellbar}) can be
generalised to non-integer arguments and is positive for $\ell \ge 2$. We can
thus use the spherical power spectrum \sout{in the Limber approximations} with
the Hankel transformation (\ref{eqn:xiGG}) to compute the two-point correlation
functions. This has the advantage of very fast numerical implementations of
the Hankel tranforms using FFTs \citep{fftlog}, which can be used in
Monte-Carlo sampling.

\mk{We compare predictions for the two-point shear correlation function using
the Hankel transformation and the full spherical Legendre transformation in
Fig.~\ref{fig:xi_p_legendre}. We show the full projection and the best
approximation, ExtL2Sph. Note that for the `FullSph' case we replace the full
projection with ExtL2Sph for $\ell > 200$ to save computation time.
%
The Hankel transform (\ref{eqn:xiGG}) is accurate at the $1$ percent level.
%
The difference between second-order Limber and full projection using the
spherical Legendre transform (\ref{eq:xi_Legendre}) is well below one percent
for all but the very large scales.}

\mk{The red circles in Fig.~\ref{fig:xi_p_legendre} present the limit of
precision we can achieve with the current fast Hankel transform implementation
for the correlation function. The green line shows the limit of the
second-order Limber approximation on the correlation function.}

>>>>

We further show that the second-order Limber approximation is an
accurate representation of the full projection, with better than percent level
precision for scales $\ell > 3$. Since this approximation involves only 1D
integrals over the matter power spectrum, it is very fast to calculate
numerically and can readily be employed in Monte-Carlo sampling methods to
obtain precision constraints on cosmological parameters.
\mk{We also compute the shear correlation function using a spherical Legendre
transformation, and compare this to the flat-sky approximated, commonly used
fast Hankel transformation.}

This paper is organised as follows.  In section~\ref{sec:wl} we provide a
pedagogical introduction to weak gravitational lensing theory, projections and
power spectra for the flat-sky and spherical cases, following the seminal work
from \citet{2000PhRvD..62d3007H} \cite[see also][]{2005PhRvD..72b3516C}. In
section~\ref{sec:L2} we derive weak-lensing observables using a second-order
Limber approximation first introduced by \citet{2008PhRvD..78l3506L}. We
compare the shear power spectrum and the commonly-used two-point shear
correlation function for the full solution to a range of different
approximations in section~\ref{sec:results}, providing cosmological constraints
for each case using CFHTLenS data from \citet{CFHTLenS-2pt-notomo}. This paper
draws from several sources of literature which have previously discussed the
full projection, or reviewed the accuracy of the Limber approximation for weak
lensing studies, namely \citet{2008PhRvD..78d3002S,2012PhRvD..86b3001B,
2012MNRAS.422.2854G, 2016arXiv161104954K}. We present a discussion and
comparison of our results to these papers in Appendix~\ref{app:B}.

@@

\mk{This type of cross-power spectrum between different cosmological epochs
$\chi$ and $\chi^\prime$ was introduced in \cite{2005PhRvD..72b3516C}. In
Sects.~\ref{sec:geom_mean} and \ref{sec:factor_D} this unequal-time
cross-spectrum \citep{2016arXiv161200770K} will be further evaluated and
simplified in the context of the Limber approximation. The oscillating Bessel
functions in equation~(\ref{eq:C_ell_phi_Pm}) ensure that only relatively close epochs
contribute to the lensing potential correlation. This make sense since observed
light rays from two galaxies at different positions on the sky that necessarily
converge at the observer today, pick up the density fluctuations at similar
times while propagating through the large-scale structure. A similar argument
has been made in \cite{BS01}: since the matter power spectrum scales with $k$
for $k \rightarrow 0$, there is decreasing power towards larger and larger
scales. In particular, the correlation of cosmic fields decreases strongly
above a coherence scale $|\chi - \chi^\prime| \gsim L_{\rm coh}$ which is significantly
smaller than the horizon scale $\chi_{\rm h}$. }

@@

The majority of cosmic shear analyses have used the predicted lensing power
spectrum approximated on a flat sky, neglecting the sky curvature. This is a
valid approach for past and current survey areas with an extent less than 10
degrees. To account for the sky curvature of the observed data, the shear
correlation functions from observed galaxy ellipticities is now routinely
computed using spherical coordinates, since projecting to a Cartesian plane has
been shown to cause significant biases of the two-point funtion on large scales
\citep{FSHK08}, \mk{and lead to spurious B-modes \citep{asgari/etal:2017}}. Here
we examine the effect of sky curvature on the theoretical models of the power
spectrum, and the effect on cosmological parameter inference (see
Sect.~\ref{sec:cfhtlens}).

@@

To separate the $k$- and $\chi, \chi^\prime$-terms in
equation~(\ref{eq:C_ell_full}), we first approximate the matter power
cross-spectrum between two distances by the geometric mean of the two involved
distances \citep{2005PhRvD..72b3516C,2016arXiv161200770K},
%
\begin{equation}
 P_{\rm m}(k; \chi, \chi^\prime) = \sqrt{ P_{\rm m}(k; \chi) P_{\rm m}(k; \chi^\prime) } \, .
  \label{eq:P_geom_mean}
\end{equation}
%
\mk{This form is justified when considering the linear power spectrum, and
follows when inserting the linearly evolving density contrast $\hat \delta(\vec
k; \chi) = D_+(\chi) \hat \delta_0(\vec k)$ into equation~(\ref{eq:p_m}), where
$\delta_0$ is the present-day linearly extrapolated density contrast, and $D_+$
the linear growth factor with $D_+(0) = 1$. This is a good approximation
also in the non-linear case as shown in \cite{2016arXiv161200770K}.}

@@

Following \cite{2008PhRvD..78l3506L} we expand to third order
%
\begin{equation}
  \lim_{\varepsilon \rightarrow 0} \int_0^\infty {\rm d} x \, {\rm e}^{-\epsilon (x - \nu)} g(x) {\rm J}_\nu(x)
  %\approx g(\nu) - \frac 1 2 g^{\prime\prime}(\nu) - \frac \nu 6 g^{\prime\prime\prime}(\nu) 
  \approx g(\nu) - \frac 1 2 \left.\frac{{\rm d}^2 g}{{\rm d} x^2}\right|_{x=\nu}
                 - \frac \nu 6 \left.\frac{{\rm d}^3 g}{{\rm d} x^3}\right|_{x=\nu},
\end{equation}
%
with $g(x) = k^{-1} f(k, \chi)$, $\chi=x/k$, and its derivatives $g^{(n)}(x) =
k^{-1-n} f^{(n)}(k, \chi)$ for a given $k$, where the derivatives of $f$
are with respect to the second argument $\chi$. \mk{In this series expansion, the
replacement of the integral with the evaluation of $g$ and its derivatives at
the maximum of the Bessel function is a good approximation of the integral if
$g$ is varying more slowly than the oscillating Bessel function.
% [Check this].
As we will show below, $f$ is a slowly varying function of the comoving
distance.} In our case the projection kernel is
%
\begin{equation}
  f(k, \chi) = \sqrt{P_{\rm m}(k; \chi)} \, a^{-1}(\chi) \chi^{-3/2} q(\chi).
  \label{eq:f_LA08}
\end{equation}
%
In the tomographic case, the index $i$ is added to $q$ and $f$.
%
Replacing both distance integrals in equation~(\ref{eq:C_ell_full_Pk}) by their
Taylor-expansions around the maxima $\nu(\ell) = \ell + 1/2$ of the two Bessel
functions, which are $k \chi$ and $k \chi^\prime$, respectively, yields
%
\begin{align}
  C^\gamma_{ij}(\ell) \approx & \, \ellbar^2(\ell, 2) \, \pref^2
    \int_0^\infty \frac{{\rm d} k}{k^3} \, k^{-2}
    \left[ f_i(k, \chi) - \frac{1}{2 k^2} f_i^{\prime\prime}(k, \chi)
      - \frac{\nu(\ell)}{6 k^3} f_i^{\prime\prime\prime}(k, \chi) + \ldots \right]
    %\nonumber \\
    %& \times
    \left[ f_j(k, \chi) - \frac{1}{2 k^2} f_j^{\prime\prime}(k, \chi)
    - \frac{\nu(\ell)}{6 k^3} f_j^{\prime\prime\prime}(k, \chi) + \ldots \right] \, .
  \label{eq:C_ell_limber2_dk}
\end{align}
%
Changing the integration to $\chi = \nu/k$ and collecting terms according to their $\nu$-dependence:
%
\begin{align}
  C^\gamma_{ij}(\ell) \approx & \, C^\gamma_{{\rm L1}, ij}(\ell) + C^\gamma_{{\rm L2}, ij}(\ell)
      %\nonumber \\
    = %&
    \frac{\ellbar^2(\ell, 2)}{\nu^4(\ell)} \, \pref^2
    \int_0^\infty {\rm d} \chi \, \chi^3 \, % P_{\rm m}\left( k \right)
    \Bigg\{
    %\left\{
    (f_i f_j)(\nu(\ell)/\chi, \chi)
    %\right.
      \nonumber \\
    &
    %\left.
     - \frac 1 {\nu^2(\ell)}
    \left[ \frac{\chi^2}{2} \left( f^{}_i f^{\prime\prime}_j + f_i^{\prime\prime} f^{}_j \right)(\nu(\ell)/\chi, \chi)
     + \frac{\chi^3}{6} \left( f^{}_i f^{\prime\prime\prime}_j + f^{\prime\prime\prime}_i f^{}_j \right)(\nu(\ell)/\chi, \chi)
    \right]
    + {\cal O}(\nu^{-4})
    %\right\}
    \Bigg\}
    \, .
  \label{eq:C_ell_limber12_dr}
\end{align}
%
The first term corresponds to the well-known first-order Limber approximation
\citep{1953ApJ...117..134L,1992ApJ...388..272K}, which is widely used in
weak gravitational lensing. We retrieve the (spherical) standard expression by
inserting back the projection kernel (equation~\ref{eq:f_LA08}),
%
\begin{align}
  C^\gamma_{{\rm L1}, ij}(\ell) = & \, \frac{\ellbar^2(\ell, 2)}{\nu^4(\ell)} \, \pref^2 \int {\rm d} \chi \frac{ q_i(\chi) q_j(\chi) }{a^2(\chi)}
  P_{\rm m}\left(\frac{\nu(\ell)}{\chi}; \chi\right) . \Label{ExtL1Sph}
  \label{eq:C_ell_limber1}
\end{align}
%
In the Limber approximation, modes between structures at different epochs do
not contribute to the single line-of-sight integration.

@@

With the definition (\ref{eq:P_k_chi_sep_s}), we factor out the function
$\tilde P_{\rm m}(k, \chi)$ from equation~(\ref{eq:f_LA08}), and we define the
separated kernel function $f_{\rm s}$ as
%
\begin{equation}
  f_{\rm s}(\chi) = D_+(\chi) a^{-1}(\chi) \chi^{-3/2} q(\chi) \, .
  \label{eq:f_LA08_s}
\end{equation}
%
The \mk{Limber equation up to second order} of the shear power spectrum can then be approximated as
%
\begin{align}
  C^\gamma_{ij}(\ell) = & \, C^\gamma_{{\rm L1}, ij}(\ell) 
    \nonumber \\
    & - \frac 1 {\nu^2(\ell)} \, \frac{\ellbar^2(\ell, 2)}{\nu^4(\ell)} \,
    \frac{\pref^2}{2}
    \int {\rm d} \chi \, \chi^{7/2} a^{-1}(\chi) D_+^{-1}(\chi) P_{\rm m}\left(\frac{\nu(\ell)}\chi; \chi\right)
    %\nonumber \\
    %& \times
    \left[ q^{}_i f^{\prime\prime}_{{\rm s} j} + f^{\prime\prime}_{{\rm s}i} q^{}_j
      + \frac{\chi}{3} \left( q^{}_i f^{\prime\prime\prime}_{{\rm s}j} + f^{\prime\prime\prime}_{{\rm s}i} q^{}_j
      \right)
    \right](\chi) \, . \Label{ExtL2Sph}
  \label{eq:C_ell_limber2_dr_s}
\end{align}
%
We compute the numerical higher derivatives as follows. The functions $f_{\rm
s}(\chi)$ are fitted as power laws with index $\approx \mk{-1.5}$, \mk{which is
expected if the growth suppression factor $D_+(a)/a$ varies slowly with $\chi$,
and the lensing efficiency $q \approx 1$ for small and medium $\chi$, given the
CFHTLenS redshift range.} The fit is carried out between $\chi_{\rm min} = 0.001
\, {\rm Mpc}/h$ and $\chi_{\rm max} = 500 \, \mbox{Mpc}/h$. At larger comoving
distances the kernel decreases faster than a power law, so we exclude this
range from the fit. Even though on those scales the derivatives are larger due
to the steeper decline, the associated errors are very small as these scales
are down-weighed by the kernel function $f_{\rm s}$ itself. At $\chi = 500 \,
\mbox{Mpc}/h$ the filter function is less than $10^{-4}$ of its value 
at $1$\,Mpc$/h$.

@@

The extended flat-sky Limber approximation is readily derived from the
spherical case equations~(\ref{eq:C_ell_limber1}, \ref{eq:C_ell_limber2_dr_s}), by
replacing the pre-factor $\ellbar^2(\ell, 2)$ with $\ell^4$,
%
\begin{align}
  P^\gamma_{{\rm L1}, ij}(\ell) = & \, p(\ell) \, \pref^2 \int {\rm d} \chi \frac{ q_i(\chi) q_j(\chi) }{a^2(\chi)}
  P_{\rm m}\left(\frac{\nu(\ell)}{\chi}; \chi\right);
  \label{eq:P_ell_limber1}
  \\
    P^\gamma_{{\rm L2}, ij}(\ell) = & - \frac 1 {\nu^2(\ell)} \, p(\ell) \, \frac{\pref^2}{2}
    \int {\rm d} \chi \, \chi^{7/2} a^{-1}(\chi) D_+^{-1}(\chi) P_{\rm m}\left(\frac{\nu(\ell)}\chi; \chi\right)
    %\nonumber \\
    %& \times
    \left[ q^{}_i f^{\prime\prime}_{{\rm s}j} + f^{\prime\prime}_{{\rm s}i} q^{}_j  
      + \frac{\chi}{3} \left( q^{}_i f^{\prime\prime\prime}_{{\rm s}j} + f^{\prime\prime\prime}_{{\rm s}i} q^{}_j
      \right)
    \right](\chi)
  \label{eq:P_ell_limber2}
\end{align}
%
Further approximations can be made for the pre-factor $p(\ell) = \ell^4/\nu^4(\ell)$ and $\nu(\ell)$:
%
\begin{enumerate}
  \item $p(\ell) = 1$, this corresponds to $\nu(\ell) = \ell$, which is the
    \emph{standard} Limber approximation \LabelTxt{L1Fl}
    Until recently, i.e.~for all pre-2014 CFHTLenS results, \ch{DES-SV \citep{abbott/etal:2016} and DLS \citep{jee/etal:2016} analyses, 
    this was the approximation of choice.}
    Note that we do not discuss the second-order Limber approximation \ch{with $p(\ell)=1$}.
  %
  \item $p(\ell) = \ell^4/(\ell + 1/2)^4$. This corresponds to the \emph{extended} Limber
    approximation \LabelTxt{ExtL1Fl, ExtL2Fl} with $\nu(\ell) = \ell + 1/2$; however the following case is typically employed:
  %
  \item $p(\ell) = 1$, but keeping as argument of the power spectrum $\nu(\ell) = \ell + 1/2$. This is
    a \emph{hybrid} between
    standard and extended Limber approximation~\LabelTxt{ExtL1FlHyb, ExtL2FlHyb}, and the first-order case 
    was used in \cite{KiDS-450,joudaki/etal:2016,joudaki/etal:2017}. As is shown
    below, this is a better approximation to the full projection than case (ii). In equation~(\ref{eq:P_ell_limber2})
    the second-order suppression factor is also left to be $\nu^{-2}(\ell) = (\ell + 1/2)^{-2}$, providing
    a slightly more accurate approximation compared to $\nu^{-2}(\ell) = \ell^{-2}$.
\end{enumerate}
%

@@

Considering first the flat-sky cases, the standard first-order Limber
approximation, (L1Fl), that was adopted for all pre-2014 CFHTLenS analyses, \ch{DES-SV and DLS analyses}, we
find \mk{it} to be accurate to better than 10\% for $\ell>3$, converging slowly to the
true projection with percent level precison at $\ell>100$. For the extended
Limber approximations `hybrid' cases (ExtL1FlHyb and ExtL2FlHyb), despite
decreased accuracy for $\ell < 8$ in comparison to the standard first-order
Limber case, the errors with respect to the true power spectrum decrease much
faster, as $\ell^{-2}$, such that percent-level precision is reached at
$\ell>15$.  The first-order extended Limber approximation `hybrid' cases
(ExtL1FlHyb) was adopted by \citet{joudaki/etal:2016}, \citet{joudaki/etal:2017}, and
\cite{KiDS-450}\footnote{We confirm that there is a typographical error in
equation 4 of \cite{KiDS-450} which should include the extra term of `+0.5' in
$\nu(\ell)$ that was incorporated in the cosmological analysis.}.

@@

On a sphere, correlation functions formally cannot be related to the power
spectrum by the Hankel transform in equation~(\ref{eqn:xiGG}), and should
be replaced by a Legendre transform \ch{@MK - ADD A REF HERE},
%
\mk{
\begin{equation}
 \xi_+(\theta) = \frac 1{4\pi} \sum\limits_{\ell=2}^\infty (2 \ell + 1) C^\gamma(\ell) {\rm P}_\ell[\cos(\theta)],
  \label{eq:xi_Legendre}
\end{equation}
%
where ${\rm P}_\ell(x)$ are the Legendre polynomials in $x$ of order $\ell$.}

The spherical power spectrum is formally not defined for non-integer $\ell$
\citep[see][for alternative spherical-sky formulae for the two-point
correlation function]{2005PhRvD..72b3516C}, as functions defined on the
sphere are necessarily periodic.
As we have shown in section~\ref{sec:comp}, however, the spherical second-order
extended Limber approximation provides a percent-level precision representation
of the full spherical projection for $\ell > 3$.
The spherical pre-factor $\ellbar(\ell, 2)$ (equation~\ref{eq:ellbar}) can be
generalised to non-integer arguments and is positive for $\ell \ge 2$. We can
thus use the spherical power spectrum \sout{in the Limber approximations} with
the Hankel transformation in equation~(\ref{eqn:xiGG}) to compute the two-point correlation
functions. This has the advantage being able to use very fast numerical implementations of
the Hankel tranforms using FFTs \citep{fftlog}, when Monte-Carlo sampling.

\mk{We compare predictions for the two-point shear correlation function using
the Hankel transformation and the full spherical Legendre transformation in
Fig.~\ref{fig:xi_p_legendre}. We show the full projection and the best
approximation, ExtL2Sph. Note that for the `FullSph' case we replace the full
projection with ExtL2Sph for $\ell > 200$ to reduce computation time.}
%
\ch{We find that the Hankel transform (equation~\ref{eqn:xiGG}) is accurate to better than $2$ percent level.
%
The difference between second-order Limber and full projection using the
spherical Legendre transform (\ref{eq:xi_Legendre}) is well below one percent
for $\vartheta < 200$ arcminutes.}
\mk{The red circles in Fig.~\ref{fig:xi_p_legendre} present the limit of
precision that we can achieve with the current fast Hankel transform implementation
for the correlation function. The green line shows the limit of the
second-order Limber approximation on the correlation function.}
